---
title: "670036975, MTHM503 Coursework, Sections B and C"
output:
  html_document: default
---

# Section B questions

```{r echo=FALSE, message=FALSE, warning=FALSE}

library(tidyverse)
library(Hmisc)
library(dplyr)

expected <- as.data.frame(read.csv("C:\\Users\\amber_000\\data_science\\mthm503\\copdexpected.csv"))
observed <- as.data.frame(read.csv("C:\\Users\\amber_000\\data_science\\mthm503\\copdobserved.csv"))

summary(observed)
observed

admissions <- sum(colSums(Filter(is.numeric, observed)))
admissions

smr <- cbind(expected[1],round(expected[-1]/observed[-1],1))
## smr_contents <- array(c(unlist(subset(smr, select = -c(Name)))))
smr_smooth <- smr

plot(smr$Name, smr$E2001)

### SMR = observed deaths / expected deaths

```
## 1

### Between 2001 and 2010, there were `r admissions` admissions to hospital in the UK for COPD

## 2

`r smr`

### Above is a visualisation of the raw standardised morbidity ratios of this data. Below is the aforementioned visualisation after the data has been smoothed.

`r smr_smooth`

### When comparing these two data visualisations, it can be noted that smoothing has had a noticeable effect on the appearance of the results.

## 3

### Over time, it appears that 

## 4

### Based on the results of the analyses of the COPD data (as outlined above), there are a few measures that ought to be prioritised if coping with COPD hospitalisations is a high priority for the NHS. This includes


# Section C questions

```{r echo=FALSE, message=FALSE, warning=FALSE}

library(cluster)
library(factoextra)
library(plyr)

characteristics <- as.data.frame(read.csv("C:\\Users\\amber_000\\data_science\\mthm503\\Characteristics.csv"))
new_substations <- as.data.frame(read.csv("C:\\Users\\amber_000\\data_science\\mthm503\\NewSubstations.csv"))

load("C:\\Users\\amber_000\\data_science\\mthm503\\January_2013.RData")

# define a data frame that contains just the columns with the substation readings

readings <- as.data.frame(subset(January_2013, select = -c(Date, Substation)))
readings$Averages <- rowMeans(January_2013[,3:145])
days <- data.frame(unique(January_2013$Date))

## Getting and scaling the average daily demands, and putting them into a single dataframe

Mean <- scale(readings$Averages)
Date <- January_2013$Date
Substation <- January_2013$Substation
average_daily_demand <- data.frame(Mean, Date, Substation)
add <- as.data.frame(aggregate(average_daily_demand$Mean, list(average_daily_demand$Substation), FUN=mean))
add

## finding the method with the highest agglomerative coefficient to use for the hierarchical cluster

m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")
ac <- function(x) {
  agnes(add, method = x)$ac
}

(sapply(m,ac))

# highest agglomerative coefficient is for the Ward method, therefore that is the one that will be used

cluster <- agnes(add, method="ward")

## dendrogram

dendrogram <- pltree(cluster, cex=0.6, hang=-1, main="Dendrogram")

## what is the optimal number of clusters

gap_statistic <- clusGap(add, FUN=hcut, nstart=25, K.max=10, B=50)
fviz_gap_stat(gap_statistic)

## based on the above, the optimal number of clusters seems is 9

## distance matrix

dist_matrix <- dist(add, method="euclidean")

final_cluster <- hclust(dist_matrix, method="ward.D2")

groups <- cutree(final_cluster, k=9)

final <- cbind(add, cluster=groups)
final

## plotting demand

demand_plots = ggplot(new_substations, aes(x=Substation) + geom_line(aes(y=) + geon_line(aes()) + geom_line(aes()) + geom_line(aes()) + geom_line(aes())))

## working with Characteristics.csv



## allocating new substations

new_substations


```

## 1

### i

#### The data in the Characteristics.csv dataset consists of a series of information about a series of power substations.   `r `

### ii

## 2

### i

#### Dendrogram: 

### ii

#### x number of clusters ended up being chosen here

### iii

## 3 

### i

#### Average daily demand plots: `r demand_plots`

### ii

### iii

## 4

### i 